---
title: "Machine Learning Project"
author: "Srikar Murali"
date: "October 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(rpart)
library(rattle)
library(randomForest)
library(MASS)

```

## Introduction

Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

Data

The training data for this project are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

Goal

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.


## Get Data Sets


Get data and partition accordingly. Do some pre-cleaning to account for NA's.
```{r get}

set.seed(31459)
train <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv',header = TRUE, na.strings = c("NA", "",'#DIV/0!'))
test <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv',header = TRUE, na.strings = c("NA", "",'#DIV/0!'))

partition <- createDataPartition(train$classe, list = FALSE, p=0.6)
trainSet <- train[partition,]
testSet <- train[-partition,]

```


## Clean Data

Remove NA's. This will reduce the size of the dataset. After this we must account for the variables with values near 0. These values will inflate the variance and offer little predictive power, thus it is important to remove them. We remove these values from both the train and test set. Also remove variables that are not needed. 
```{r clean}

trainSet <- trainSet[,colSums(is.na(trainSet)) == 0]
testSet <- testSet[,colSums(is.na(testSet)) == 0]

nearzerovar <- nearZeroVar(trainSet)
trainSet <- trainSet[, -nearzerovar]
testSet <- testSet[, -nearzerovar]

nas <-  sapply(trainSet, function(x) mean(is.na(x))) > 0.95
trainSet <- trainSet[, nas==FALSE]
testSet <- testSet[, nas==FALSE]

trainSet <- trainSet[, -(1:5)]
testSet <- testSet[, -(1:5)]


```


## Build Models

Start building the model. To partition the dataset, we will use k-fold cross validation. The first model will be built using the random forest approach. Next I will try and use the rpart approach. I will then use the naive bayes approach. Lastly I will use the gbm method.
```{r build}

ctrl <- trainControl(method = 'cv', number = 3, verboseIter = FALSE)
fit <- train(classe ~ ., preprocess = c("center", "scale"), data = trainSet, method = 'rf', trControl = ctrl)
fit$finalModel

#fit2 <-train(classe ~ ., preprocess = c("center", "scale"), data = trainSet, method = 'rpart', trControl = ctrl)
#fit3 <- train(classe ~., data = trainSet, method = "nb")
#fit4 <- train(classe ~ ., data=trainSet, method = "gbm",
#              trControl = ctrl,
#              verbose = FALSE)
predictions <- predict(fit, newdata = testSet)
predictions <- predict(fit, newdata = testSet)
confusionMatrix(testSet$classe, predictions)
plot(predictions, color = 'red', ylab = 'count', xlab = 'category', title = 'Classification')

```

The random forest gave the best classification accuracy. Thus, I will use them. The other algorithims take a very long time to run, thus I will comment them out for now. However feel free to uncomment if you wish to test them.

## Retrain model

To ensure accuracy, we must rain the model on the full training set. Thus, we must make the test and training set the same as the ones that were just tested.
```{r tune}
nearzv <- nearZeroVar(train)
train <- train[, -nearzv]
test <- test[, -nzv]

nas2 <- sapply(train, function(x) mean(is.na(x))) > 0.95
train <- train[, nas2 == FALSE]
test <- test[, nas2 == FALSE]

train <- train[, -(1:5)]
test <- test[, -(1:5)]

ctrl2 <- trainControl(method = 'cv', number = 3, verboseIter = FALSE)
fitFin <- train(classe ~., data = train, method = 'rf', trControl = ctrl2)

pml_write_files <- function(x) {
  n <- length(x)
  for(i in 1:n) {
    filename <- paste0("problem_id_", i, ".txt")
    write.table(x[i], file=filename, quote=F, row.names=F, col.names=F)
  }
}

pml_write_files(as.character(predictions))

```

The random forest model gave the best estimation of the error

Predictions : B A B A A E D B A A B C B A E E A B B B